---
title: "Final Project"
author: "Christopher Butler"
date: "May 2, 2016"
output: html_document
---
For this Final Project I wanted to duplicate and further investigate a question about video game console popularity and states. Xbox vs PlayStation: Which Does America Love More? http://www.decluttr.com/blog/2016/03/18/xbox-vs-playstation-xbox-one-ps4/
I wanted to go deeper, using Google’s Adsense Keyword Planner tool to grab data at the county level. I also wanted to see if there were any interesting correlations with Demographics like Population density, Average Income, Unemployment, etc. It would be very interesting to see if console preferences tracked with Income level, or some other demographic indicator. 

To do this I had to scrape the Google Keyword Planner.  Apparently Google does not open their Adsense API because of abuses with SEOs in the past. It takes months to get approved to use it, and even then you need to have an existing project for them to investigate. 
I ended up using iMacro and creating scripts to download the information into CSV files for further ingestion and manipulation. 

For the demographic information, I wanted to use MIT’s latest data experiment DATAUSA http://datausa.io/. DATAUSA does have a completely open API for downloading data using a REST interface. You can even get CSV formatted data. I went with the CSV formatted data to save time, instead of learning the jsonr package. 

I had a lot of data wrangling to do even before getting into R. I had to create a script to combine all of the keyword planner CSV files. After that I had to combine that data into a format that can be easily used in R. I opted for a google doc, CSV export. 

I had issues with the DATAUSA data not having good county names to join with the Keyword Planner Data. I lucked up and found an excellent Google provided resource called Fusion Tables. It had most county information in table form. I scraped it and put it into CSV files for consumption. The Fusiontables information and data is located here https://support.google.com/fusiontables/answer/1182141?hl=en


Data Wrangler Helpers
Keyword Planner Data:
https://docs.google.com/spreadsheets/d/1VVIprSPTFlmTAafUaPvjHQKJvA5nJM9dBExCd096cmw/pub?gid=0&single=true&output=csv

DataUSA:
http://api.datausa.io/api/csv/?show=geo&sumlevel=county&year=latest&required=pop

State Abbrivations:
https://docs.google.com/spreadsheets/d/1Utgyxd_nBvaaQM3wczd4zyA2B5RrhAhxgKXxbzuavSo/pub?gid=0&single=true&output=csv

FusionTable Data:
https://docs.google.com/spreadsheets/d/1hENyGRj4nSXYJkhDipAw-BOMGBrcAQK8FT_kewEI818/pub?gid=0&single=true&output=csv


```{r}
library(readr)
library(dplyr)
library(rjson)
library(stringr)
library(tidyr)

keyword_data <- read_csv("https://docs.google.com/spreadsheets/d/1VVIprSPTFlmTAafUaPvjHQKJvA5nJM9dBExCd096cmw/pub?gid=0&single=true&output=csv")
keyword_update <- keyword_data
keyword_update$county_key <- str_replace_all(keyword_update$county_key, "new york", "new york,")
keyword_update <- keyword_update %>% mutate( county_key =  ifelse(!str_detect(county_key,","),str_replace_all(county_key," ", ","), county_key))
keyword_update <- keyword_update %>% 
  mutate( console_type =  ifelse(!str_detect(keyword,"x"),"Playstation", "Xbox")) 
keyword_update <- keyword_update %>%  separate(county_key, into = c("state", "county_name"), sep = "\\,") 
keyword_update <- keyword_update %>%  mutate(county_name = str_trim(  str_replace_all(county_name, ",", " ")))




counties_pop <- read_csv("http://api.datausa.io/api/csv/?show=geo&sumlevel=county&year=latest&required=pop")
counties_pop <- counties_pop %>%  separate(geo_name, into = c("county_old", "state"), sep = "\\,") 
counties_pop <- counties_pop %>% mutate(state = str_trim(state))


#counties_pop <- fromJSON(file="http://api.datausa.io/api/?show=geo&sumlevel=county&year=latest&required=pop")

state_abbr <- read_csv("https://docs.google.com/spreadsheets/d/1Utgyxd_nBvaaQM3wczd4zyA2B5RrhAhxgKXxbzuavSo/pub?gid=0&single=true&output=csv")
state_abbr <- state_abbr %>% mutate(state=str_to_lower(State), Abbreviation=str_to_lower(Abbreviation))
keyword_update <- keyword_update %>%  left_join(state_abbr)

geoid_key <- read_csv("https://docs.google.com/spreadsheets/d/1hENyGRj4nSXYJkhDipAw-BOMGBrcAQK8FT_kewEI818/pub?gid=0&single=true&output=csv")
#counties_pop <- counties_pop %>% left_join(geoid_key, by=c("geo" ="GEO_ID"))
counties_pop <- counties_pop %>% separate(geo, into= c("geo_prefix","geo_num"), sep="US") %>% mutate(geo_num = as.numeric(geo_num))
geoid_key <- geoid_key %>% mutate(GEO_ID2 = as.numeric(GEO_ID2 ))
counties_pop <- geoid_key %>% left_join(counties_pop, by=c("GEO_ID2"="geo_num"))

counties_pop <- counties_pop %>% mutate(county_name = str_to_lower(`County Name`), state=str_to_lower(state))



#keyword_update <- keyword_update %>%  left_join(state_abbr)
keyword_update <- keyword_update %>% left_join(counties_pop, by = c("Abbreviation"="state abbr", "county_name" = "county_name"))



```

After importing the data, I started looking for interesting correlations

```{r, echo=FALSE}
keywords_combined <- keyword_update %>% 
  group_by(State,county_name,console_type, pop) %>% 
  summarize(total_searches = sum(average))


keyword_totals <- keyword_update %>% 
  group_by(State,county_name,console_type, pop) %>% 
  summarize(total_searches = sum(average)) %>% filter(!is.na(pop))

testp <- keyword_totals %>% filter(console_type == "Playstation")
testx <- keyword_totals %>% filter(console_type == "Xbox")



plot(testp$pop, testp$total_searches, col="red",pch=15)
points(testx$pop, testx$total_searches, col="green", pch=16)
myline <- lm(testx$total_searches ~ testx$pop)
summary(myline)
abline(myline)
```

